Investigating Determinism of Gaming Engines for Autonomous Vehicle Verification

### 2019-12-20 paper/2_Background.tex
.....
Did some minor tweaks

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{Other/Figures/GameEngineLoop.pdf}
\caption{Game engine loop block diagram~\cite{GameProgPatternsBook}}
\label{GameEngineLoopDiagram}
\end{figure}
The first part of the game loop is to process the user inputs which may take the form of user keyed inputs or, in the case of CAV development, the resultant actions of the autonomous vehicle given the current state of the environment. For example, the throttle and steering inputs from the vehicle controller of an AV
%
The update interprets the intended actions of all dynamic actors in the scene, performs any necessary physics calculations and updates the actor states. 
%
The rendering engine then takes the updated actor states and renders the scene~\cite{GameProgPatternsBook}. 
% 
It is necessary to have an understanding of how the game loop operates in order to understand potential sources of non-determinism that are discussed below. 
% 
The key aspects of the game loop needed for this discussion are as follows, see Fig.~\ref{GameEngineLoopDiagram}:

@@

\begin{figure*}[b]
    \centering
    \includegraphics[width=0.99\linewidth]{Other/Figures/MethodologyDiagram.pdf}
    \caption{Shows a flow diagram of the methodology proposed to define the level of simulation variance of a gaming engine for usage in CAV simulations.}

>>>>

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{Other/Figures/GameEngineLoop.pdf}
\caption{Game engine loop block diagram~\cite{GameProgPatternsBook}}
\label{GameEngineLoopDiagram}
\end{figure}
The first part of the game loop is to process the user inputs which may take the form of user keyed inputs or, in the case of CAV development, the resultant actions of the autonomous vehicle given the current state of the environment. For example, the throttle and steering inputs from the vehicle controller of an AV.
%
The update interprets the intended actions of all dynamic actors in the scene, performs any necessary physics calculations and updates the actor states. 
%
The rendering engine then takes the updated actor states and renders the scene~\cite{GameProgPatternsBook}. 
% 
It is necessary to have an understanding of how the game loop operates in order to understand potential sources of non-determinism that are discussed below. 
% 
The key aspects of the game loop needed for this discussion are as follows, see Fig.~\ref{GameEngineLoopDiagram}:

@@

\begin{figure*}[b]
    \centering
    \includegraphics[width=0.99\linewidth]{Other/Figures/MethodologyDiagram.pdf}
    \caption{Shows a flow diagram of the methodology proposed to define the simulation variance of a gaming engine for usage in CAV simulations.}

### 2019-12-20 paper/6_Conclusions.tex
Did some minor tweaks

\section{Conclusions}\label{s:conclusion}
We have provided a summary of the sources of non-determinisim in game engines. This was followed by presenting a method by which one can determine the level of simulation variance of a gaming engine used in V\&V application of CAV simulations. The method consists of five stages, starting with designing the experiment, then suggesting the various settings that can improve or worsen the performance, and finally focuses on running the experiment and analysing the output.

A case study is then used to show the implementation of the method introduced, where several tests were setup and run to explore the different interactions between the different agents in a simulator called CARLA. It was found that for the computer used (with specs of an i9 9960X CPU chip, NVIDIA GeForce RTX 2080 graphics card and 64GB of RAM) the level of non determinism is sufficiently low to assume the simulator to be deterministic as long as the CPU + GPU utilisation is below 50\% and that tests are terminated as soon as a collision of any type is detected. %since it was noticed that collisions cause the level of non determinism to jump remarkable invalidating the assumption of the simulator is deterministic.

>>>>

\section{Conclusions}\label{s:conclusion}
We have provided a summary of the sources of non-determinisim in game engines. This was followed by presenting a method by which one can determine the simulation variance of a gaming engine used in V\&V application of CAV simulations. The method consists of five stages, starting with designing the experiment, then suggesting the various settings that can improve or worsen the performance, and finally focuses on running the experiment and analysing the output.

A case study is then used to show the implementation of the method introduced, where several tests were setup and run to explore the different interactions between the different agents in a simulator called CARLA. It was found that for the computer used (with specs of an i9 9960X CPU chip, NVIDIA GeForce RTX 2080 graphics card and 64GB of RAM) the simulation variance was sufficiently low to assume the simulator to be deterministic as long as the CPU + GPU utilisation is below 50\% and that tests are terminated as soon as a collision of any type is detected. %since it was noticed that collisions cause the level of non determinism to jump remarkable invalidating the assumption of the simulator is deterministic.
